Demes: a standard format for demographic models

### 2022-05-30 paper.tex

By allowing the definition of selfing and cloning probabilities, we allow many
standard models to be defined.
However, by parameterising selfing and cloning as we have, we assume that these
properties of populations can be specified independently from the genetics.
In other words, mutations that cause selfing probabilities to fluctuate within
an epoch are not considered.
More details of the mathematical properties of selfing and cloning rates
in a coalescent context can be found in
\citet{nordborg1997coalescent,hartfield2016facsexcoal}.

@@

\subsection{Static models, not parameterised models}
\label{sec:appendix-static}

@@

Similarly, one might be interested in running simulations in which
the demographic model parameters are drawn from a distribution, e.g.,
in ABC inference~\citep{beaumont2002approximate}.
Other inference procedures based on optimising a loss function
\citep{gutenkunst2009inferring,kamm2017efficient,jouganous2017inferring,ragsdale2019models,excoffier2021fastsimcoal2}
%(\ggcomment{more refs? MCMC? neural networks?})
need users to specify parameter bounds,
and possibly non-linear or conditional constraints between parameters.
Indeed, the choice of how to parameterize a model could be important for
some inference methods (e.g. absolute times versus relative times between events).

@@

The model is, by design, very limited. We do not specify details of
individuals or of the processes that happen within populations,
because such details are necessarily complicated, vary greatly by
application, and attempts to encompass such a broad range of biological
processes are unlikely to succeed.
The narrow focus on individuals and their groupings into populations
should ensure that the Demes standard is stable, and not subject to
continual development as more and more elaborations of basic processes
are added. Demes is intended to provide \emph{part} of a
simulation model specification; other parts of the specification,
such as sampling strategies, or parameters that vary by time
or population can \emph{refer to} this well defined and unambigous
description of the demography.

>>>>

By allowing the definition of selfing and cloning probabilities, we allow many
standard models to be defined.
However, by parameterizing selfing and cloning as we have, we assume that these
properties of populations can be specified independently from the genetics.
In other words, mutations that cause selfing probabilities to fluctuate within
an epoch are not considered.
More details of the mathematical properties of selfing and cloning rates
in a coalescent context can be found in
\citet{nordborg1997coalescent,hartfield2016facsexcoal}.

@@

\subsection{Static models, not parameterized models}
\label{sec:appendix-static}

@@

Similarly, one might be interested in running simulations in which
the demographic model parameters are drawn from a distribution, e.g.,
in ABC inference~\citep{beaumont2002approximate}.
Other inference procedures based on optimizing a loss function
\citep{gutenkunst2009inferring,kamm2017efficient,jouganous2017inferring,ragsdale2019models,excoffier2021fastsimcoal2}
%(\ggcomment{more refs? MCMC? neural networks?})
need users to specify parameter bounds,
and possibly non-linear or conditional constraints between parameters.
Indeed, the choice of how to parameterize a model could be important for
some inference methods (e.g. absolute times versus relative times between events).

@@

The model is, by design, very limited. We do not specify details of
individuals or of the processes that happen within populations,
because such details are necessarily complicated, vary greatly by
application, and attempts to encompass such a broad range of biological
processes are unlikely to succeed.
The narrow focus on individuals and their groupings into populations
should ensure that the Demes standard is stable, and not subject to
continual development as more and more elaborations of basic processes
are added. Demes is intended to provide \emph{part} of a
simulation model specification; other parts of the specification,
such as sampling strategies, or parameters that vary by time
or population can \emph{refer to} this well defined and unambiguous
description of the demography.
